The role of continuous developmental change in explaining discontinuities in word learning

### 2019-01-29 cogsci/cogsci.Rmd
second draft cogsci

---
title: "Continuous Development in Learning Similar-sounding Words"

@@

header-includes:
   - \usepackage{tipa}

@@

    "Stager and Werker (1997) first showed that children's early ability to distinguish similar sounds do not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using simpler testing methods) that children do encode such contrasts, thus favoring a rather a continuous pattern of development. The present study proposes a precise probabilistic model describing how development may proceed in a continuous fashion across the lifespan. The purpose of the model is to account for previously documented facts while providing new predictions.  We collected data from both preschool children and adults, and we showed that the model can explain various patterns of learning both within the same age and across development. Our work highlights the role of computational modeling in advancing our understanding of development both through organizing exsinting knowledge and through generating new principled hypotheses."

@@

Over the first year of life, children become sensitive to the phonetic variations that are used to distinguish meanings in their native language [@werker1984]. One could imagine that these perceptual skills would be automatically applied to the task of word learning. However, developmental data show that 14 m.o children find it challenging to associate minimally different (but perceptually discriminated) sounds such as "bin" and "din" to different objects [@stager1997].

Several factors can explain this finding. For example it is possible that the task of meaning learning increases cognitive demands on children (compared to a simple perceptual discrimination). In particular, it requires paying attention to both the sounds and the corresponding objects, which may hinder precise encoding in memory of some phonetic details [@stager1997; @hofer2017]. Additional difficulty might arise from ambiguous phonological boundaries at this stage of development [e.g., @rost2009], or from uncertainty about the referential status of the novel word [@fennell2010].

Regardless of the exact explanation, it is generally accepted that by around 17 m.o, children succeed in the same task [@werker2002]. What could be the mechanism of development? Early accounts assumed–whether implicitly or explicitly– that children encode words in a binary way: they either fail or succeed in encoding the relevant phonetic details (simultaneously with the meanings). This account suggested a discontinuous/stage-like pattern of development whereby younger children fail to encode the contrastive phonetic detail, whereas older children succeed.

Nevertheless, subsequent findings suggest otherwise. On the one hand, 14-month-olds–who typically fail in the original task–succeed when an easier *testing* method is used, even under the same *learning* conditions [@yoshida2009]. They also succeed when uncertainty is mitigated via disambiguating cues [e.g., @thiessen2007]. On the other hand, adults show patterns of learning similar to those shown by 14-month-olds when the task is more challenging and when similarity between words increases [@white2013; @pajak2016].

These evidence point towards another scenario, where the representations are encoded in a probabilistic (rather than binary) way, and where development is continuous, rather than stage-like [@swingley2007; @yoshida2009; @pajak2016]. On this account, correct representations are learned early in development, but these representations are encoded with higher uncertainty in younger children, leading to apparent failure in relatively demanding tasks. Development is a continuous process whereby the initial noisy representations become more precise. Crucially, more precise representation are still imperfect: Even with adults show low accuracy learning when the sounds are subtle enough, e.g., non-native sounds [@pajak2016].

We provide an intuitive illustration of how such an account explains patterns of learning and development in Figure \ref{fig:illus}. We observe low accuracy in word learning when the perceptual distance between the labels is small relative to the uncertainty/tolerance with which these labels are encoded. For example, in Stager and Werker's original experiment, children are supposed to associate label 1 ("bih") and label 2 ("dih") with object 1 and object 2, respectively.  Though children could learn that the label "bih" is a better match to object 1 than "dih", they could still judge the sound "dih" as a plausible instance of the label "bih", thanks to the relatively large tolerance of the encoding.

Accuracy in word learning increases when the perceived distance between the labels is large relative to the tolerance of their encoding. This improvement can occur in the same developmental stage if the perceptual distance between labels is enhanced through using different-sounding labels (e.g., "lif" vs. "neem" instead of "bih" and "dih") or through using additional cues [@thiessen2007; @rost2009]. Besides, accuracy can improve across development through the refinement of the encoding.

Building on this intuition, the current work proposes a precise probabilistic model, which we use to both account for previous experimental findings, and to make new predictions that have not been tested before. Using new data collected from both preschool children and adults, we show that the model can explain various patterns of learning both within the same age and across development.

@@

illustration <- data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "low accuracy", type='Disambig.') %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.5, s), rep = "2", case = "low accuracy", type='Disambig.')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "high accuracy", type='Disambig.')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=1.1, Vr=0.5, s), rep = "2", case = "high accuracy", type='Disambig.')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "low accuracy", type='Develop.')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.5, s), rep = "2", case = "low accuracy", type='Develop.')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.1, s), rep = "1", case = "high accuracy", type='Develop.')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.1, s), rep = "2", case = "high accuracy", type='Develop.'))

illustration$case <- factor(illustration$case, levels = c("low accuracy", "high accuracy"))
illustration$type <- factor(illustration$type, levels = c("Disambig.", "Develop."))

@@

illus_plot <- ggplot(illustration, aes(x=x, y=y, linetype=rep)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=9),
      legend.text=element_text(size=9),
      axis.text = element_text(size = 9),
      strip.text = element_text(size = 9),
      axis.title = element_text(size = 9),
      aspect.ratio = 0.5
      )+

  geom_col(aes(x = x, y = y, fill = rep),
               position=position_dodge(), 
               width=0.2,
               data = subset(illustration, x==0))+
  geom_text(aes(label=paste("Acc=", prob, sep="")),
            x=1.7, y=1, size=2.5, 
            data=probs)+
  xlab("Perceptual space") + ylab("Prob") +
  facet_grid(type~case) +
  theme(legend.position="none")

@@

```{r illus, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:illus} An illustration of the probabilistic/continuous account using simulated data. Two words are represented as distributions over the perceptual space. Here accuracy expresses the extent to which a given sound instance indicates the word on the left (red bar), as opposed to the word on the right (green bar). The values vary between 0.5 (total overlap between the distributions) and 1 (no overlap). Accuracy is low when the perceptual distance between labels is small relative to the variance. Accuracy increases when the perceptual distance is enhanced (through disambiguation), or when the variance decreases (e.g., through development)."}

illus_plot
```

@@

## Task

We model the word learning task introduced by @stager1997, and a testing method similar to the one used by @yoshida2009. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. After this exposure phase, participants perform a series of two-alternative forced choices. In each testing trial, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \ref{fig:task}.

@@

@@

## Simulations

<!--If we denote the correct/target object as $o_T$ and the incorrect object as $o_{\lnot T}$, then the probability of chosing the correct object is: $P(o_T|s_T) = $ -->

During the testing phase (see Figure \ref{fig:task}), participants are presented with one target sound $s_T \in\{s_1, s_2\}$ and with both objects $o_1$ and $o_2$. In order to make a choice, we need to determine which object is more probable under the target sound $s_T$. In other words, we need to compare the probabilities $P(o_1|s_T)$ and $P(o_2|s_T)$. The values of these probabilities can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$
The joint probability $P(o, C, L, s)$ is obtained by factoring the Bayesian network in Figure \ref{fig:model}:
$$P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) $$
which could be transformed using Bayes rule into:

$$
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
$$

From this general expression we derive the exact analytical formula, which expresses the probability of accurate responses in the testing phase of the task:

\begin{equation}
P(o_T|s_T)= \frac{1 + e^{-(\Delta s^2 + \Delta o^2) /2\sigma^2}}{1 + e^{-(\Delta s^2 + \Delta o^2) /2\sigma^2} + e^{-\Delta s^2 /2\sigma^2} + e^{-\Delta o^2 /2\sigma^2 }}
\end{equation}

```{r model, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:model}Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters."}

@@

In order to have a more quantitative understanding of the model, we simulate the values of the predicted accuracy (Expression 1) as a function of the perceptual distance between the sounds $\Delta s$. We used as parameters the two remaining variables, i.e., the visual distance between the semantic referents $\Delta o$ and the standard deviation of the distributions $p(s| L)$ and $p(o | C)$ (which, in this simulation, are assumed to have similar values, i.e., $\sigma =\sigma_C \approx \sigma_L$). The simulations are shown in Figure \ref{fig:simulation}.
```{r }

#Import data from saved
simulations <- feather::read_feather("../saved_data/simulations.feather")

plot_simulation <- ggplot(simulations, aes(x=x, y=y, col=sigma)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11),
      aspect.ratio = 0.7
      )+
  #plot.margin=grid::unit(c(0,0,0,0), "mm")
  ylim(c(0.4,1))+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("sound distance") + ylab("Prob. accurate") +
  facet_grid(.~object_distance)

#ggsave("figs/data_all.png", plot = plot_data_all, width = 7, height = 3)

```{r simulation, fig.env = "figure*", fig.pos = "h", fig.width=7, fig.height=3, fig.align = "center", strip.white=TRUE, fig.cap = "\\label{fig:simulation}The predicted probability of accurate responses in the testing phase, as a function the perceptual distance between the sounds. Colors indicate different values of the standard deviation which we assume is common to both the labels' and concepts' probabilistic representations. Panels represent graphs using different values of the visual distance between the objects."}
plot_simulation
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

The simulations explain some previousely documented facts, and make new predictions:

@@

3) For fixed values of $\Delta s$ and $\sigma$, accuracy increases with the visual distance between the semantic referents $\Delta o$. This is a new prediction that our model makes. Previous work studied the effect of several bottom-up and top-down properties in disambiguating similar sounding words [@rost2009; @thiessen2007; @dautriche2015; @fennell2010], but no previous study tested the effect of the visual distance between the semantic referents.

To sum, we introduced a model that accounts for some qualitative learning patterns observed in previous studies, and makes a new prediction. In the experiment below, we test whether the model makes accurate *quantitative* predictions by fitting it to new experimental data collected from preschool children and adults.

@@

In this experiment, we tested participants in the word learning task introduced above (Figure \ref{fig:task}). We explored all three parameters of the model. In particular, both the sound similarity ($\Delta s$) and object similarity ($\Delta o$) were varied simultaneously in a within-subject design. Besides, two age groups (preschool children and adults) were tested on the same task to explore whether development can be characterized with the uncertainty parameter, $\sigma$, in the probabilsitic representations.
<!--Children were trained to learn the association between pairs of nonesense words (e.g., "lif"/"neem") and pairs of objects (e.g., two flowers). After training, children performed a series of two-alternative forced choices. In each testing trial, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives (e.g., flower 1 or flower 2). Each child learned an artificial lexicon where both the pairs of sounds and the pairs of objects can be similar or different. Our predictions were as follows. If the sound and meaning interact in word learning, then the pair of words that differ on both the meaning and sound levels should be the easiest to learn, followed by the pairs of words that are different on one level, but similar on another level. Finally, the pair of words that are similar on both levels should be the most difficult to learn.  -->

@@

```{r}
sim_transform <- feather::read_feather("../saved_data/d_sim_transform.feather")

sim_plot <- ggplot(sim_transform, 
      aes(x = labels, y = normalized_all, group = factor(stimuli), col=factor(stimuli))) +
  geom_point(size=2)+
  geom_line()+
  #geom_pointrange(aes(ymin = lower, ymax = upper), 
     #            position = position_dodge(width = .1)) + 
  scale_colour_discrete(name="Stimuli")+
  theme_few()+
theme(legend.title = element_text(size=7.5),
      legend.text=element_text(size=7.5),
      axis.text = element_text(size = 7.5),
      strip.text = element_text(size = 7.5),
      axis.title = element_text(size = 7.5),
      aspect.ratio = 0.7
      ) +
  ylim(c(0,1.1)) +
  xlab("Stimuli group")+ ylab("Perceptual distance (experimental)")

@@

```{r stim, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:stim}Normalized distances for both sound and object pairs used in this study."}

@@

#Import data from saved
data_human_model <- feather::read_feather("../saved_data/data_human_model.feather") %>%
  mutate(source = ifelse(key == "human", "Experimental",
                         ifelse(key == "prediction", "Model 1", "Model 2")))

@@

plot_data <- ggplot(data_human_model, 
#ggplot(subset(model_data_plot, group=='children'),
      aes(x = sound_dist, y = value, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) + facet_grid(group~source, scales="free_y") +
  theme(aspect.ratio = 0.7)

#ggsave("figs/data_all.png", plot = plot_data_all, width = 7, height = 3)

@@

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.width=7, fig.height=3, fig.align = "center", strip.white=TRUE, fig.cap = "\\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). Experimental results are shown on the left. Predictions from Model 1 (one free parameter) and Model 2 (two free parameters) are shown in the middle and on the right, respectively."}
plot_data
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

Preschoolers were asked if they would be willing to play a game on a tablet with the experimenter and were informed that they could stop playing at any time. The experimenter explained that the game consisted in learning some words spoken in an alien planet. The experiment began with two simple examples (not included in the analysis), and in these examples children were given feedback from the experimenter so as to make sure they correctly understood the structure of the task. After the introduction and examples, children were tested in a sequence of five conditions: the four experimental conditions plus the catch condition. In each condition, participants saw a first block of four exposure trials followed by four testing trials, and a second block of two exposure trials (for memory refreshment) followed by an additional four testing trials.

In the exposure trials, children saw two objects associated with their corresponding sounds. We presented the first object on the left side of the tablet's screen simultaneously with the corresponding sound. The second sound-object association followed on the other side of the screen after 500ms. For both objects, visual stimuli were present for the duration of the sound clip (∼800ms).  In the testing trials, children saw both objects simultaneously and heard only one sound. They completed the trial by selecting which of the two objects corresponded to the sound. They responded by touching one of the pictures on the tablet.

The object-sound pairings were randomized across participants, as was the order of the conditions (except for the catch condition which was always placed in the middle of the testing sequence). We also randomized the on-screen position (left vs. right) of the two pictures on each testing trial.

The procedure in the adult experiment was identical to that of preschool children, except that preschoolers were accompanied by an experimenter and used a tablet, whereas adults used their local computers to complete the experiment online.

### Model fitting

>>>>

---
title: "The role of continuous developmental change in explaining discontinuities in word learning"

@@

header-includes:
   - \usepackage{tipa}
   - \usepackage[rightcaption]{sidecap}

@@

    "Cognitive development is often characterized in term of discontinuities. Nevertheless, in some cases, this characterization can be problematic and unnecessary. In this paper, we show that a phenomenon that appears stage-like, can be well described by a continuous account.  We used as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using simpler testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a precise probabilistic model describing how development may proceed in a continuous fashion across the lifespan.  The model accounts for previously documented facts and provides new predictions.  We collected data from preschool children and adults, and we showed that the model can explain various patterns of learning both within the same age and across development. The findings suggest that other aspects of cognitive development which are typically thought of as discontinuities, may emerge from simpler, continuous mechanisms."

@@

Cognitive development is sometimes characterized in terms of a succession of discontinuous stages [@piaget1954]. Such characterization may be problematic as it often renders the study of the learning mechanism quite challenging. Indeed, it is unclear how a precise algorithm––whose steps are gradual and logically related––can account for transitions between stages involving qualitatively different representations [e.g., @carey2009].

Nevertheless, at least in some cases, development may only appear to be stage-like. This appearance can be due, for example, to the use of a cognitively-demanding task which may mask learning, or to the use of hard statistical thresholding (in particular, p-value < 0.05) which creates a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, as we shall see in this paper, a continuous model––involving similar representations across the lifespan––may provide a simpler and more transparent account of development.

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds ("dih" vs. "bih") does not automatically translate into word learning skills (i.e., associating these sounds with different meanings). Subsequent research has focused on proposing possible explanations for this observed gap between speech perception and word learning [e.g., @fennell2010; @stager1997; @hofer2017;  @rost2009].

In this work we focus, rather, on the development of this phenomenon. In fact, regardless of the exact explanation, it is generally accepted that by around 17 m.o, children succeed in the same task [@werker2002]. How does development proceed? Early accounts assumed that children encode words in a binary way: they either fail or succeed in encoding the relevant phonetic details (simultaneously with the meanings). This account suggested a discontinuous/stage-like pattern of development whereby younger children fail to encode the contrastive phonetic detail, whereas older children succeed.

<!--Over the first year of life, children become sensitive to the phonetic variations that are used to distinguish meanings in their native language [@werker1984]. One could imagine that these perceptual skills would be automatically applied to the task of word learning. However, developmental data show that 14 m.o children find it challenging to associate minimally different (but perceptually discriminated) sounds such as "bih" and "dih" to different objects [@stager1997].  -->

<!--can explain this finding. For example it is possible that the task of meaning learning increases cognitive demands on children (compared to a simple perceptual discrimination). In particular, it requires paying attention to both the sounds and the corresponding objects, which may hinder precise encoding in memory of some phonetic details [@stager1997; @hofer2017]. Additional difficulty might arise from ambiguous phonological boundaries at this stage of development [e.g., @rost2009], or from uncertainty about the referential status of the novel word [@fennell2010]. -->

Nevertheless, subsequent findings have suggested otherwise. On the one hand, 14-month-olds–who typically fail in the original task–succeed when an easier testing method is used, even under the same learning conditions [@yoshida2009]. They also succeed when uncertainty is mitigated via disambiguating cues [e.g., @thiessen2007]. On the other hand, adults show patterns of learning similar to those shown by 14-month-olds when the task is more challenging and when similarity between words increases [@white2013; @pajak2016].

These evidence point towards another scenario, where the representations are encoded in a probabilistic (rather than binary) way, and where development is continuous, rather than stage-like [@swingley2007; @yoshida2009; @pajak2016]. On this account, correct representations are learned early in development, but these representations are encoded with higher uncertainty in younger children, leading to apparent failure in relatively demanding tasks. Development is a continuous process whereby the initial noisy representations become more precise. Crucially, more precise representation are still imperfect: Even adults show low accuracy learning when the sounds are subtle enough, e.g., non-native sounds [@pajak2016].

We provide an intuitive illustration of how such an account explains patterns of learning and development in Figure \ref{fig:illus}. We observe low accuracy in word learning when the perceptual distance between the labels is small relative to the uncertainty with which these labels are encoded. For example, in Stager and Werker's original experiment, children are supposed to associate label 1 ("bih") and label 2 ("dih") with object 1 and object 2, respectively.  Though children could learn that the label "bih" is a better match to object 1 than "dih", they could still judge the sound "dih" as a plausible instance of the label "bih", thanks to the relatively large uncertainty of the encoding.

According to this account, accuracy in word learning improves if we increase either the perceptual distinctiveness of the stimuli (e.g., through using different-sounding labels), or the precision of the encoding itself (e.g., across development). Building on this intuition, the current work proposes a precise probabilistic model, which we use to both account for previous experimental findings, and to make new predictions that have not been tested before. Using new data collected from both preschool children and adults, we show that the model can explain various patterns of learning both within the same age and across development.

<!-- "lif" vs. "neem" instead of "bih" and "dih" -->

@@

illustration <- data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "low accuracy", type='Distinctiveness') %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.5, s), rep = "2", case = "low accuracy", type='Distinctiveness')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "high accuracy", type='Distinctiveness')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=1.1, Vr=0.5, s), rep = "2", case = "high accuracy", type='Distinctiveness')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "low accuracy", type='Precision')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.5, s), rep = "2", case = "low accuracy", type='Precision')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.1, s), rep = "1", case = "high accuracy", type='Precision')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.1, s), rep = "2", case = "high accuracy", type='Precision'))

illustration$case <- factor(illustration$case, levels = c("low accuracy", "high accuracy"))
illustration$type <- factor(illustration$type, levels = c("Distinctiveness", "Precision"))

@@

@@

```{r fig.width=2, fig.height=2.5}
illus_plot <- ggplot(illustration, aes(x=x, y=y, col=rep)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=9),
      legend.text=element_text(size=9),
      axis.text = element_text(size = 9),
      strip.text = element_text(size = 9),
      axis.title = element_text(size = 9),
      aspect.ratio = 0.7
      )+

  geom_col(aes(x = x, y = y, fill = rep),
               position=position_dodge(), 
               width=0.2,
               data = subset(illustration, x==0))+
  geom_text(aes(label=paste("Acc=", prob, sep="")),
            x=1.7, y=1, size=3, 
            data=probs, 
            col='black')+
  xlab("Perceptual space") + ylab("Prob") +
  facet_grid(type~case) +
  theme(legend.position="none")

```{r illus, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:illus} An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or green). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy increases when the stimuli are less similar (left panel), or when the representation are more precise (right panel)."}

img <- png::readPNG("figs/illustration.png")
grid::grid.raster(img)
```

@@

@@

\begin{SCfigure}
\centering
\includegraphics[width=1.5in]{figs/model.png}
\caption{Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters.}
\label{fig:model}
\end{SCfigure}

## Inference
<!--If we denote the correct/target object as $o_T$ and the incorrect object as $o_{\lnot T}$, then the probability of chosing the correct object is: $P(o_T|s_T) = $ 
(see Figure \ref{fig:task})
-->
The learner hears a sound $s$ and has to decided which object $o$ provides an optimal match to this sound. To this end, the learner has to compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$
The joint probability $P(o, C, L, s_T)$ is obtained by factoring the Bayesian network in Figure \ref{fig:model}:
$$P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) $$
which could be transformed using Bayes rule into:

\begin{equation}
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\end{equation}

<!--
```{r model, fig.env = "figure", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:model}Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters."}

```
-->

## Task and model predictions

@@

We use the model to predict performance in the word learning task introduced by @stager1997, and a testing method similar to the one used by @yoshida2009. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \ref{fig:task}.

We use the general Equation 1 to derive the exact analytical expression for the probability of accurate responses $p(o_T | s)$ (target object $o_T$ given a sound $s$) in the simple case of two-alternative forced choice in the testing phase of our experimental task:

\begin{equation}
P(o_T|s)= \frac{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}}{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}+ e^{-\Delta s^2 /2\sigma_L^2} + e^{-\Delta o^2 /2\sigma_C^2 }}
\end{equation}

Figure \ref{fig:simulation} show simulations of the predicted accuracy (Expression 2) as a function of the distinctiveness parameters ($\Delta s$ and $\Delta o$) and the precision parameter, i.e., the variance of the distributions $p(s| L)$ and $p(o | C)$ (which, for ease of the simulations' interpretation, are assumed to have similar values, i.e., $\sigma =\sigma_C \approx \sigma_L$).

```{r, echo=FALSE}
pdf.options(encoding='CP1250')
```

#Import data from saved
simulations <- feather::read_feather("../saved_data/simulations.feather") %>%
  mutate(obj = ifelse(object_distance == "0.25", "\u0394o = 0.25",
                      ifelse(object_distance == "0.5","\u0394o = 0.5",
                             "\u0394o = 1")))

plot_simulation <- ggplot(simulations, aes(x=x, y=y, col=sigma)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 8),
      aspect.ratio = 0.7
      )+
  #plot.margin=grid::unit(c(0,0,0,0), "mm")
  ylim(c(0.4,1))+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab(expression(Delta*"s")) + ylab(expression("P("*o['T']*"| s)")) +
  facet_grid(.~obj) +
  #facet_grid(.~object_distance) +
  scale_color_discrete(name = expression(sigma))

ggsave('simulations.png', plot = plot_simulation, width = 7, height = 2)
#ggsave('simulations.png', plot = plot_simulation)
```

```{r simulation, fig.env = "figure*", fig.pos = "h", fig.width=7, fig.height=3, fig.align = "center", strip.white=TRUE, sanitize=FALSE, fig.cap = "\\label{fig:simulation}The predicted probability of accurate responses in the testing phase as a function of stimuli distinctiveness $\\Delta s$ (x-axis) and $\\Delta o$ (different panels) and representation precision $\\sigma$ (color). Dashed line represents chance."}
#plot_simulation
#ggsave(plot_simulation, filename = 'simulations.png')
img <- png::readPNG("simulations.png")
grid::grid.raster(img)

@@

The simulations explain some previously documented facts, and make new predictions:

@@

3) For fixed values of $\Delta s$ and $\sigma$, accuracy increases with the visual distance between the semantic referents $\Delta o$. This is a new prediction that our model makes. Previous work studied the effect of several bottom-up and top-down properties in disambiguating similar sounding words [e.g., @fennell2010; @rost2009; @thiessen2007], but no previous study tested the effect of the visual distance between the semantic referents.

In sum, we introduced a model that accounts for some qualitative learning patterns observed in previous studies, and makes a new prediction. In the experiment below, we test whether the model makes accurate quantitative predictions by fitting it to new experimental data collected from preschool children and adults.

@@

In this experiment, we tested participants in the word learning task introduced above (Figure \ref{fig:task}). We explored all three parameters of the model. In particular, both the sound similarity ($\Delta s$) and object similarity ($\Delta o$) were varied simultaneously in a within-subject design. Two age groups (preschool children and adults) were tested on the same task to explore whether development can be characterized with the uncertainty parameter, $\sigma$, in the probabilistic representations. The Experiment and the predictions were pre-registered.
<!--Children were trained to learn the association between pairs of nonesense words (e.g., "lif"/"neem") and pairs of objects (e.g., two flowers). After training, children performed a series of two-alternative forced choices. In each testing trial, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives (e.g., flower 1 or flower 2). Each child learned an artificial lexicon where both the pairs of sounds and the pairs of objects can be similar or different. Our predictions were as follows. If the sound and meaning interact in word learning, then the pair of words that differ on both the meaning and sound levels should be the easiest to learn, followed by the pairs of words that are different on one level, but similar on another level. Finally, the pair of words that are similar on both levels should be the most difficult to learn.  -->

@@

```{r}
sim_transform <- feather::read_feather("../saved_data/d_sim_process.feather")

sim_plot <- ggplot(sim_transform, 
      aes(x = labels, y = mean, group = factor(stimuli), col=factor(stimuli))) +
  #geom_point(size=2)+
  geom_line()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  scale_colour_discrete(name="Stimuli")+
  theme_few()+
theme(legend.title = element_text(size=7.5),
      legend.text=element_text(size=7.5),
      axis.text = element_text(size = 7.5),
      strip.text = element_text(size = 7.5),
      axis.title = element_text(size = 7.5),
      aspect.ratio = 0.7
      ) +
  ylim(c(0,1.1)) +
  xlab("Stimuli group")+ ylab("Perceptual distance (experimental)")

@@

```{r stim, fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=2.5, fig.cap = "\\label{fig:stim}Distances for both sound and object pairs from an adult norming study. Error bars represent 95\\% confidence intervals."}

@@

#Import data from saved
data_human_model <- feather::read_feather("../saved_data/data_human_model.feather") %>%
  mutate(source = ifelse(key == "human", "Experimental",
                         ifelse(key == "prediction", "Model 1", "Model"))) %>%
  filter(source != "Model 1")

@@

plot_data <- ggplot(data_human_model, 
#ggplot(subset(model_data_plot, group=='children'),
      aes(x = sound_dist, y = value, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      #) + facet_grid(group~source, scales="free_y") +
      ) + facet_grid(source~group) +
  theme(aspect.ratio = 0.7)

#ggsave("figs/data_all.png", plot = plot_data, width = 3, height = 6)
#ggsave("figs/data_all.png", plot = plot_data, width = 8, height = 16)

@@

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=12, fig.cap = "\\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (top) and model predictions (bottom). Error bars represent 95\\% confidence intervals."}
plot_data
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

<!--and were informed that they could stop playing at any time. The experimenter explained that the game consisted in learning some words spoken in an alien planet.

The experiment began with two simple examples (not included in the analysis), and in these examples children were given feedback from the experimenter so as to make sure they correctly understood the structure of the task. After the introduction and examples,
-->

Preschoolers were tested at the nursery school using a tablet, whereas adults used their local computers to complete the same experiment online. Participants were tested in a sequence of five conditions: the four experimental conditions plus the catch condition. In each condition, participants saw a first block of four exposure trials followed by four testing trials, and a second block of two exposure trials (for memory refreshment) followed by an additional four testing trials.

In the exposure trials, participants saw two objects associated with their corresponding sounds. We presented the first object on the left side of the tablet's screen simultaneously with the corresponding sound. The second sound-object association followed on the other side of the screen after 500ms. For both objects, visual stimuli were present for the duration of the sound clip (∼800ms).  In the testing trials, participants saw both objects simultaneously and heard only one sound. They completed the trial by selecting which of the two objects corresponded to the sound. The object-sound pairings were randomized across participants, as was the order of the conditions (except for the catch condition which was always placed in the middle of the testing sequence). We also randomized the on-screen position (left vs. right) of the two pictures on each testing trial.
